---
title: "Overview of mdmaps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of mdmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  out.width = "100%"
)
```

```{r setup}
library(mdmaps)
library(ggplot2)
library(sf)
```

`mdmaps` ships pre-projected `sf` objects and companion tibbles for
making maps of Maryland. Every `sf` object is in **EPSG:26985**
(NAD83 / Maryland, meters) so layers can be combined without
reprojection.

## The base map

The state outline and counties are the two layers you'll reach for most
often.

```{r counties}
ggplot() +
  geom_sf(data = md_state_sf, fill = "grey95", color = "grey40") +
  geom_sf(data = md_counties_sf, fill = NA, color = "grey20", linewidth = 0.3) +
  theme_void()
```

## Political districts

Maryland has 8 U.S. House districts, 47 state legislative districts
(each electing 1 Senator), and 71 House of Delegates electing units
(some whole-LD multi-member, some A/B/C sub-districts).

```{r districts}
ggplot() +
  geom_sf(data = md_state_sf, fill = "grey95") +
  geom_sf(data = md_congressional_districts_sf,
          aes(fill = district_num), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "Maryland congressional districts") +
  theme_void()
```

## Census geographies

Three decennial vintages of tracts ship, plus 2020 blocks, both PUMA
vintages, and ZCTAs with 2019 ACS population.

```{r zctas}
ggplot(md_zcta_sf) +
  geom_sf(aes(fill = population), color = NA) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90",
                       name = "Population") +
  labs(title = "Population by ZCTA (2019 ACS)") +
  theme_void()
```

## Municipalities

Maryland is not a strong-municipality state — most suburban land is
unincorporated county territory. `md_municipalities_sf` gives the 157
or so incorporated places.

```{r munis}
ggplot() +
  geom_sf(data = md_counties_sf, fill = "grey95", color = "grey60") +
  geom_sf(data = md_municipalities_sf, fill = "steelblue", color = NA) +
  theme_void()
```

## MDOT SHA districts

State Highway Administration engineering districts — the boundary
SHA uses for maintenance and CHART incident response.

```{r sha}
ggplot(md_sha_districts_sf) +
  geom_sf(aes(fill = factor(district_num)), color = "white") +
  scale_fill_brewer(palette = "Set2", name = "District") +
  theme_void()
```

## Lookups

Companion tibbles round out the spatial layers:

```{r lookups}
head(md_counties)
head(md_zips)
```
